### 2017-05-11 Publications/Original Software Publications/Science of Computer Programming/others/cover letter/Cover-Letter.tex

\documentclass[11pt]{article}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{url}

%\usepackage{mathtools}
%\mathtoolsset{showonlyrefs}

\usepackage{color}
\usepackage{natbib}
% \usepackage[style=authoryear-comp,dashed=false]{biblatex}
% \bibliography{piecewise_lyapunov}

\textwidth 15cm
\setlength{\textheight}{1.1\textheight}
\newcommand\hi{\hspace*{\parindent}}
\newcommand\vi{\vspace{\baselineskip}}
\newcommand\lac{{\mbox{{\Huge\bf L}\hspace{-0.65em}
\raisebox{-1.2ex}{\Huge\bf A}\hspace{-1.1em}
\raisebox{-0.6ex}{\Huge\bf C~}}}}

\newcommand{\Real}{\mathbb{R}}

\pagestyle{empty}

\begin{document}

\hoffset -1.5cm \voffset .5cm

\sf

\vspace*{-2.5cm} \hspace*{-0.1cm} {\small
{\mbox{\begin{minipage}{1.5cm}
\centerline{\includegraphics[width=1.8cm]{logo_ufam.png}}
\end{minipage}}} \hspace*{0.1cm}
\begin{minipage}{11.0cm}{\sf \large \textbf{FEDERAL UNIVERSITY OF AMAZONAS}}\\
{\sc Department of Electronics and Computing}
\end{minipage}
} \vspace*{0mm}

\hspace*{-.7cm} {\rule[-1ex]{15cm}{0.03cm}}

\vi \vi
\begin{flushright}
\begin{minipage}{7.0cm}\small
{\bf Please Reply to:}\\
Mr. Felipe R. Monteiro\\
Federal University of Amazonas\\
Department of Electronics and Computing\\
Avenida General Rodrigo Ot\'avio, 6200, Coroado\\
69077-000 - Manaus - AM - BRAZIL\\
{\em e-mail: felipemonteiro@ufam.edu.br}
\end{minipage}
\end{flushright}

\vi \vi
\hspace*{\fill}{\small Manaus, \today.}
\vi \vi

\begin{flushleft}
    Professor Mark van den Brand\\
    Editor-in-Chief\\
    Science of Computer Programming Software Track
    \end{flushleft}
\vi

\indent Dear Editor-in-Chief,
\vi

Please find enclosed our manuscript entitled as ``{\em ESBMC-GPU: A Context-Bounded Model Checking Tool to Verify CUDA Programs}'' by Felipe R. Monteiro, Erickson H. Alves, Isabela S. Silva, Hussama A. Ismail, Lucas C. Cordeiro, and Eddie B. de Lima Filho, which we would like to submit for publication in the software track of the journal of {\bf Science of Computer Programming}.

\vi

This paper addresses ESBMC-GPU's functionality and discusses its practical capabilities and applications, whose underlying techniques have been previously published at the Journal of Concurrency and Computation: Practice and Experience. We focus here on the usage aspects of ESBMC-GPU, in order to emphasize its design and implementation from the user point of view. In addition, we present a rigorous experimental evaluation of our approach against other state-of-the-art software verifiers ({\it i.e.}, CIVL, GKLEE, GPUVerify) using standard CUDA benchmarks taken from the literature.

\vi

For your reference, we attach the original paper (ccpe2016.pdf) with this submission. These new results have not been published elsewhere and are not under consideration by another journal.

\vi

\begin{quote}
\begin{flushright}
Thank you for your time.~~~~~~~

Sincerely,~~~~~~~

Felipe R. Monteiro~~~~~~~~
\end{flushright}
\end{quote}

\end{document}